The state learner -- a super learner for right-censored data

### 2025-05-28 manuscript/manuscript-one.tex
rewritten intro

@@

[More motivation]

A super learner is a machine learning algorithm that combines
a finite set of learners into a meta learner by estimating
prediction performance in hold-out samples using a
pre-specified loss function \citep{van2007super}. When the aim
is to make a prediction model, super learners combine strong
learners, such as Cox regression models and random survival
forests \citep[][Section 8.4]{gerds2021medical}. While the
general idea of combining strong learners based on
cross-validation stems from earlier work
\citep{wolpert1992stacked,breiman1996stacked}, the name super
learner is justified by an oracle inequality
\citep{van2003unicv,vaart2006oracle}.

We define the state learner, a new super learner for
right-censored data, which simultaneously estimates the
expected loss of learners of the event time distribution and
the censoring distribution. Compared to the super learners of
\cite{polley2011-sl-cens} and \cite{golmakani2020super}, the
state learner proposed here poses no restrictions on the type
of learners that can be included in the library. Compared to
super learners that rely on inverse probability of censoring
weighting, censoring unbiased transformations, or
pseudo-values, the state learner does not require a
pre-specified censoring distribution. Compared to the super
learners by \cite{han2021inverse} and
\cite{westling2021inference}, the state learner has
theoretical guarantees which these methods do not have.

The loss function underlying the state learner operates on
coarsened data. In the right-censored survival setting the
coarsened data consist of the minimum and the order of the
censoring time and the event time as well as the baseline
covariates. The state learner works by simultaneously
assessing how well learners of the event time distribution and
the censoring distribution predict the states of the coarsened
data. To analyse the theoretical properties of the state
learner we focus on the discrete super learner which combines
the library of learners by picking the one that minimises the
cross-validated loss \citep{van2007super}. In the presence of
competing risks, our algorithm uses separate libraries of
learners for the cumulative hazard functions for each of the
competing risks and for the censoring distribution. We show
that the oracle selector of the state learner is consistent if
all libraries contain a consistent learner and prove a finite
sample oracle inequality.

The state learner algorithm can output a medical risk
prediction model \citep{gerds2021medical} which predicts the
probability of an event based on covariates in the presence of
competing risks. Another important application, which
motivated our contribution, is in targeted learning where
conditional event probabilities occur as high-dimensional
nuisance parameters which need to be estimated at a certain
rate \citep{van2011targeted, rytgaard2021estimation,
  rytgaard2022targeted}. The state learner is particularly
well-suited for this application as it simultaneously provides
estimates for both the outcome and the censoring, both of
which are needed for targeted learning. For the asymptotic
bias term of a targeted estimator, which uses the state
learner to estimate nuisance parameters, we show that a second
order product structure holds. We illustrate both applications
of the state learner with prostate cancer data
\citep{kattan2000pretreatment}.

There are many existing super learners for right-censored
data. We briefly review them here. Machine learning based on
right-censored data commonly uses the partial log-likelihood
as a loss function
\citep[e.g.,][]{li2016regularized,yao2017deep,lee2018deephit,katzman2018deepsurv,gensheimer2019scalable,lee2021boosted,kvamme2021continuous}.
This loss function is also suggested for super learning with
right-censored data by \cite{polley2011-sl-cens}, where data
is assumed to be observed in discrete time. However, the
partial log-likelihood loss does not work well with data
splitting (cross-validation) in continuous time. The reason is
that the partial log-likelihood loss assigns an infinite value
for any learner that predicts piece-wise constant cumulative
hazard functions when there are event times in the test set
which do not occur in the learning set. This problem occurs
with prominent survival learners including the Kaplan-Meier
estimator, the random survival forest, and the semi-parametric
Cox regression model and these learners cannot be included in
the library of the super learner proposed by
\cite{polley2011-sl-cens}. When a proportional hazards model
is assumed, the baseline hazard function can be profiled out
of the likelihood \citep{cox1972regression}. The
cross-validated partial log-likelihood loss
\citep{verweij1993cross} has therefore been suggested as a
loss function for super learning by \cite{golmakani2020super}.
This choice of loss function restricts the library of learners
to include only Cox proportional hazards models, and hence
excludes many learners such as, e.g., random survival forests,
additive hazards models, and accelerated failure time models.
Other proposals for super learning with right-censored data
use an inverse probability of censoring weighted (IPCW) loss
function
\citep{graf1999assessment,van2003unicv,molinaro2004tree,keles2004asymptotically,hothorn2006survival,gerds2006consistent,gonzalez2021stacked},
censoring unbiased transformations
\citep{fan1996local,steingrimsson2019censoring}, or
pseudo-values
\citep{andersen2003generalised,mogensen2013random,sachs2019ensemble}.
All these methods rely on an estimator of the censoring
distribution, and their drawback is that this estimator has to
be pre-specified. An approach which avoids a pre-specified
censoring model was proposed independently by
\cite{han2021inverse} and \cite{westling2021inference}. In
both articles, the authors suggest to iterate between learning
of the outcome model and learning of the censoring model using
IPCW loss functions. However, it seems that this procedure is
in general not guaranteed to converge to the true
data-generating mechanism
\citep[][Appendix~A.4]{munch2024thesis}.

We introduce our notation and framework in
Section~\ref{sec:framework}. In
Section~\ref{sec:super-learning} we define general super
learning for right-censored data.
Section~\ref{sec:super-learner-simple} introduces the state
learner, and Section~\ref{sec:theor-results-prop} provides
theoretical guarantees. We report results of our numerical
experiments in Section~\ref{sec:numer-exper} and analyse a
prostate cancer data set in Section~\ref{sec:real-data-appl}.
Section~\ref{sec:discussion} contains a discussion of the
merits and limitations of our proposal. Proofs are given in
Appendix~\ref{sec:proofs}. An implementation of the state
learner is available at
\url{https://github.com/amnudn/statelearner} along with code
for reproducing our numerical experiments.

@@

As input to a super learner we need a data set
\( \data_n=\{O_i\}_{i=1}^n \) of i.i.d.\ observations from
\( P \in \mathcal{P} \) and a collection of candidate learners
$\mathcal{A}$. Each learner \(a \in \mathcal{A}\) is a map
\( a \colon \mathcal{O}^n \rightarrow \Theta \) which takes a
data set as input and returns an estimate
$a(\data_n) \in \Theta$ of $\theta$.
% Formally, the domain of $a$ depends on \(n\) but we suppress this in the notation.
In what follows, we use the short-hand notation
\(P[f] = \int f(o) P(\diff o) \). A super learner evaluates the
performance of \(a \in \mathcal{A}\) with a loss function
\(L\colon \Theta \times \mathcal{O} \rightarrow \R_+\) and
estimates the expected loss \(P[L(a(\data_n), \blank)]\) using
cross-validation. Specifically, the expected loss of $a\in\mathcal A$
is estimated by splitting the data set $\data_n$ into $K$ disjoint
approximately equally sized subsets
\(\data_n^1, \data_n^2, \dots, \data_n^K \) and then calculating the
cross-validated loss
\begin{equation*}
  % \label{eq:cv-risk-est}
  \hat{R}_n(a; L) =
  \frac{1}{K}\sum_{k=1}^{K}
  % \empmeas^k{[L {(a{ (\data_n^{-k})} , \blank) }]},
  \frac{1}{| \data_n^{k} |}\sum_{O_i \in \data_n^{k}}
  L
  {
    \left(
      a{ (\data_n^{-k})}
      , O_i
    \right)
  },
  \quad \text{with} \quad
  \data_n^{-k} = \data_n \setminus \data_n^{k}.
\end{equation*}
% where \( \empmeas^{k} \) is the empirical measure of \( \data_n^{k} \). 
The subset \(\data_n^{-k}\) is referred to as the \(k\)'th training
sample, while \(\data_n^{k}\) is referred to as the \(k\)'th test or
hold-out sample.
The discrete super learner is defined as
\begin{equation*}
\hat{a}_n = \argmin_{a\in\mathcal A}\hat{R}_n(a; L).
\end{equation*}
The oracle learner is defined as the learner that minimises the
expected loss under the data-generating distribution \( P \),
i.e.,
\begin{equation*}
  \tilde{a}_n =
  \argmin_{a \in \mathcal{A}}
  \tilde{R}_n(a; L),
  \quad \text{with} \quad 
  \tilde{R}_n(a; L)=
  \frac{1}{K}\sum_{k=1}^{K} 
  P{
    \left[
      L
      {
        \left(
          a{ (\data_n^{-k})}
          , \blank
        \right)
      }
    \right]}
  .
\end{equation*}
Note that both the discrete super learner and the oracle learner
depend on the library of learners and on the number of folds \(K\),
and that the oracle learner is a function of the data and the unknown
data-generating distribution. However, these dependencies are
suppressed in the notation.

@@

The state learner estimates the parameter \( F \) which is a
feature of the observed right-censored data distribution
\( P \). In particular, \( F \) depends on the censoring
distribution and is typically not of direct interest in
itself. In the following, we describe how an estimate of
\( F \), as provided by the state learner, can be used to
estimate more relevant parameters of interest, such as
survival probabilities.

@@

We here provide explicit formulas for obtaining event-free
survival probabilities and risk predictions given the
state-occupation function \( F \) defined in
equation~\eqref{eq:F-def}. These formula holds under the
assumption of conditional independent censoring and
positivity, as introduced in Section~\ref{sec:framework}.

By equations~(\ref{eq:lambdaj}) and~(\ref{eq:gamma}) and the
definition of \( F \), we have
\begin{equation}
  \label{eq:7}
  % \Gamma(t , x) = \int_0^t \frac{F(\diff s, -1, x )}{F(s-,
  % 0, x )}, \quad \text{and} \quad
  \Lambda_j(t , x) 
  = \int_0^t  \frac{F(\diff s, j, x )}{F(s-, 0, x )},
  \quad j \in \{1,2\}.
\end{equation}
Equation~\eqref{eq:surv-def} provides a formula for obtaining
the event-free survival probabilities from the cause-specific
cumulative hazard functions. Similarly, cause-specific risk
predictions can be obtained from \( \Lambda_1 \) and
$\Lambda_2$ using the formula
\citep[e.g.,][]{benichou1990estimates,
  ozenne2017riskregression},
\begin{equation}
  \label{eq:cs-risk-def}
  Q(T \leq t, D = j \mid X=x) =
  \int_0^t \exp\left\{-\Lambda_{1}(u \mid x)-\Lambda_{2}(u
    \mid x)\right\}  \Lambda_j(\diff u \mid x),
  \quad j \in \{1,2\}.
\end{equation}
Hence, given the state learner's estimate of \( F \) we can
use equation~\eqref{eq:7} to obtain estimates of the
cause-specific cumulative hazard functions $\Lambda_j$, which
can in turn be used to obtain estimates of the event-free
survival probabilites and cause-specific risks through
equations~\eqref{eq:surv-def} and~\eqref{eq:cs-risk-def}.

@@

... TL + SL = strong... The state learner is particularly well
suited for targeted and debiased machine learning as it
immediately provides estimates of the nuisance parameters
pertaining to the outcome and the censoring models. While it
was our original motivation for developing the state learning,
we leave the study of the state learner in the context of
targeted and debiased machine learning for a future paper.

@@

For all super learners we use a library consisting of three learners:
The Kaplan-Meier estimator
\citep{kaplan1958nonparametric,Gerds_2019prodlim}, a Cox model with
main effects \citep{cox1972regression, survival-package}, and a random
survival forest \citep{ishwaran2008random,rfsrc-paclage}. We use the
same library to learn the outcome distribution and the censoring
distribution. Note that the three learners in our library of learners
can be used to learn the cumulative hazard functions of the 
outcome and the censoring distribution. The latter works by
training the learner on the data set \( \data_n^c \), where
\( \data_n^c = \{O_i^c\}_{i=1}^n \) with
\( O_i^c = (\tilde{T}_i, 1-\Delta, X_i) \). When we say that we use a
learner for the cumulative hazard function of the outcome to learn the
cumulative hazard function of the censoring time, we mean that the
learner is trained on \( \data_n^c \).

>>>>

\usepackage[author=]{fixme}
\fxusetheme{color}
\definecolor{fxtarget}{rgb}{.5,.5,.5}
\definecolor{fxnote}{rgb}{.5,.5,.5}

\fxsetup{status=draft}

@@

Accurate risk prediction from time-to-event data is a central
topic in many areas of medical research and public health.
Super learning \citep{van2007super}, also known as ensemble
learning or stacked regression
\citep{wolpert1992stacked,breiman1996stacked}, provides a
powerful approach by combining multiple candidate prediction
models. In the context of survival analysis, a super learner
can for instance combine Cox regression models and random
survival forests \citep[][Section 8.4]{gerds2021medical}.

In this paper we propose the state learner, a new super
learner designed to handle the specific challenges of ensemble
learning with right-censored data. The state learner estimates
the performance of a library of candidate models for both the
event-time and censoring distributions simultaneously. This
joint estimation framework avoids the need for pre-specifying
a censoring model, which is a limitation of many existing
methods. Our approach also allows for full flexibility in the
choice of models in the library, contrasting with earlier
proposals that restrict the library to specific model classes
\citep{polley2011-sl-cens,golmakani2020super}. Furthermore,
unlike iterative approaches such as those in
\cite{han2021inverse} and \cite{westling2021inference}, we can
provide theoretical guarantees for the performance of the
state learner.

The state learner is based on the simple idea of considering
right-censored competing risks data as an artificial
multi-state system where censoring is included as a state of
its own. Candidate models are then assessed based on how well
they predict the state occupations of this system across time,
based on baseline covariates. We demonstrate how to construct
a library for the state learner using existing survival
models, and how to obtain risk predictions from the resulting
ensemble. To analyse the theoretical properties of the state
learner we focus on the discrete super learner which pick the
model in the library with best estimated performance
\citep{van2007super}. We show that the discrete state learner
is consistent under mild conditions and prove a finite-sample
oracle inequality.

\subsection{Related work}
\label{sec:related-work}

There are many existing super learners for right-censored
data. Machine learning based on right-censored data commonly
uses the partial log-likelihood as a loss function
\citep[e.g.,][]{li2016regularized,yao2017deep,lee2018deephit,katzman2018deepsurv,gensheimer2019scalable,lee2021boosted,kvamme2021continuous}.
This loss function is also suggested for super learning with
right-censored data by \cite{polley2011-sl-cens}, where data
is assumed to be observed in discrete time. However, the
partial log-likelihood loss does not work well with data
splitting (cross-validation) in continuous time. The reason is
that the partial log-likelihood loss assigns an infinite value
for any learner that predicts piece-wise constant cumulative
hazard functions when there are event times in the test set
which do not occur in the learning set. This problem occurs
with prominent survival learners including the Kaplan-Meier
estimator, the random survival forest, and the semi-parametric
Cox regression model and these learners cannot be included in
the library of the super learner proposed by
\cite{polley2011-sl-cens}. When a proportional hazards model
is assumed, the baseline hazard function can be profiled out
of the likelihood \citep{cox1972regression}. The
cross-validated partial log-likelihood loss
\citep{verweij1993cross} has therefore been suggested as a
loss function for super learning by \cite{golmakani2020super}.
This choice of loss function restricts the library of learners
to include only Cox proportional hazards models, and hence
excludes many learners such as, e.g., random survival forests,
additive hazards models, and accelerated failure time models.

Alternative approaches for super learning with right-censored
data use an inverse probability of censoring weighted (IPCW)
loss function
\citep{graf1999assessment,van2003unicv,molinaro2004tree,keles2004asymptotically,hothorn2006survival,gerds2006consistent,gonzalez2021stacked},
censoring unbiased transformations
\citep{fan1996local,steingrimsson2019censoring}, or
pseudo-values
\citep{andersen2003generalised,mogensen2013random,sachs2019ensemble}.
All these methods rely on an estimator of the censoring
distribution, and their drawback is that this estimator has to
be pre-specified.

More recent work by \cite{han2021inverse} and
\cite{westling2021inference} circumvents the need to
pre-specify a censoring model by iterating
between estimation of the outcome and censoring models. However, it seems that this procedure is in general not
guaranteed to converge to the true data-generating mechanism
\citep[][Appendix~A.4]{munch2024thesis}.

\subsection{Outline of the paper}
\label{sec:outline-paper}

We introduce our notation and framework in
Section~\ref{sec:framework}. Section~\ref{sec:super-learning}
defines general super learning for right-censored data.
Section~\ref{sec:super-learner-simple} presents the state
learner. In Section~\ref{sec:use-cases-state}, we demonstrate
how the state learner can be used to obtain risk predictions
and how it can be applied in the context of causal inference.
Section~\ref{sec:theor-results-prop} provides theoretical
guarantees. Section~\ref{sec:numer-exper} reports the results
of numerical experiments, and Section~\ref{sec:real-data-appl}
illustrates the method on prostate cancer data. We conclude
with a discussion in Section~\ref{sec:discussion}. Proofs are
included in Appendix~\ref{sec:proofs}. Code and an
implementation of the state learner are available at
\url{https://github.com/amnudn/statelearner}.

% A super learner is a machine learning algorithm that combines
% a finite set of learners into a meta learner by estimating
% prediction performance in hold-out samples using a
% pre-specified loss function \citep{van2007super}. When the aim
% is to make a prediction model, super learners combine strong
% learners, such as Cox regression models and random survival
% forests \citep[][Section 8.4]{gerds2021medical}. While the
% general idea of combining strong learners based on
% cross-validation stems from earlier work
% \citep{wolpert1992stacked,breiman1996stacked}, the name super
% learner is justified by an oracle inequality
% \citep{van2003unicv,vaart2006oracle}.

% We define the state learner, a new super learner for
% right-censored data, which simultaneously estimates the
% expected loss of learners of the event time distribution and
% the censoring distribution. Compared to the super learners of
% \cite{polley2011-sl-cens} and \cite{golmakani2020super}, the
% state learner proposed here poses no restrictions on the type
% of learners that can be included in the library. Compared to
% super learners that rely on inverse probability of censoring
% weighting, censoring unbiased transformations, or
% pseudo-values, the state learner does not require a
% pre-specified censoring distribution. Compared to the super
% learners by \cite{han2021inverse} and
% \cite{westling2021inference}, the state learner has
% theoretical guarantees which these methods do not have.

% The loss function underlying the state learner operates on
% coarsened data. In the right-censored survival setting the
% coarsened data consist of the minimum and the order of the
% censoring time and the event time as well as the baseline
% covariates. The state learner works by simultaneously
% assessing how well learners of the event time distribution and
% the censoring distribution predict the states of the coarsened
% data. To analyse the theoretical properties of the state
% learner we focus on the discrete super learner which combines
% the library of learners by picking the one that minimises the
% cross-validated loss \citep{van2007super}. In the presence of
% competing risks, our algorithm uses separate libraries of
% learners for the cumulative hazard functions for each of the
% competing risks and for the censoring distribution. We show
% that the oracle selector of the state learner is consistent if
% all libraries contain a consistent learner and prove a finite
% sample oracle inequality.

% The state learner algorithm can output a medical risk
% prediction model \citep{gerds2021medical} which predicts the
% probability of an event based on covariates in the presence of
% competing risks. Another important application, which
% motivated our contribution, is in targeted learning where
% conditional event probabilities occur as high-dimensional
% nuisance parameters which need to be estimated at a certain
% rate \citep{van2011targeted, rytgaard2021estimation,
%   rytgaard2022targeted}. The state learner is particularly
% well-suited for this application as it simultaneously provides
% estimates for both the outcome and the censoring, both of
% which are needed for targeted learning. For the asymptotic
% bias term of a targeted estimator, which uses the state
% learner to estimate nuisance parameters, we show that a second
% order product structure holds. We illustrate both applications
% of the state learner with prostate cancer data
% \citep{kattan2000pretreatment}.

@@

As input to a super learner we need a data set
\( \data_n=\{O_i\}_{i=1}^n \) of i.i.d.\ observations from
\( P \in \mathcal{P} \) and a collection of candidate models
or learners $\mathcal{A}$. Each learner \(a \in \mathcal{A}\)
is a map \( a \colon \mathcal{O}^n \rightarrow \Theta \) which
takes a data set as input and returns an estimate
$a(\data_n) \in \Theta$ of $\theta$.
% Formally, the domain of $a$ depends on \(n\) but we suppress this in the notation.
In what follows, we use the short-hand notation
\(P[f] = \int f(o) P(\diff o) \). A super learner evaluates the
performance of \(a \in \mathcal{A}\) with a loss function
\(L\colon \Theta \times \mathcal{O} \rightarrow \R_+\) and
estimates the expected loss \(P[L(a(\data_n), \blank)]\) using
cross-validation. Specifically, the expected loss of $a\in\mathcal A$
is estimated by splitting the data set $\data_n$ into $K$ disjoint
approximately equally sized subsets
\(\data_n^1, \data_n^2, \dots, \data_n^K \) and then calculating the
cross-validated loss
\begin{equation*}
  % \label{eq:cv-risk-est}
  \hat{R}_n(a; L) =
  \frac{1}{K}\sum_{k=1}^{K}
  % \empmeas^k{[L {(a{ (\data_n^{-k})} , \blank) }]},
  \frac{1}{| \data_n^{k} |}\sum_{O_i \in \data_n^{k}}
  L
  {
    \left(
      a{ (\data_n^{-k})}
      , O_i
    \right)
  },
  \quad \text{with} \quad
  \data_n^{-k} = \data_n \setminus \data_n^{k}.
\end{equation*}
% where \( \empmeas^{k} \) is the empirical measure of \( \data_n^{k} \). 
The subset \(\data_n^{-k}\) is referred to as the \(k\)'th training
sample, while \(\data_n^{k}\) is referred to as the \(k\)'th test or
hold-out sample.
The discrete super learner is defined as
\begin{equation*}
\hat{a}_n = \argmin_{a\in\mathcal A}\hat{R}_n(a; L).
\end{equation*}
The oracle learner is defined as the learner that minimises the
expected loss under the data-generating distribution \( P \),
i.e.,
\begin{equation*}
  \tilde{a}_n =
  \argmin_{a \in \mathcal{A}}
  \tilde{R}_n(a; L),
  \quad \text{with} \quad 
  \tilde{R}_n(a; L)=
  \frac{1}{K}\sum_{k=1}^{K} 
  P{
    \left[
      L
      {
        \left(
          a{ (\data_n^{-k})}
          , \blank
        \right)
      }
    \right]}
  .
\end{equation*}
Note that both the discrete super learner and the oracle learner
depend on the library of learners and on the number of folds \(K\),
and that the oracle learner is a function of the data and the unknown
data-generating distribution. However, these dependencies are
suppressed in the notation.

@@

The state learner estimates the parameter \( F \) which is a
feature of the observed right-censored data distribution
\( P \). In particular, \( F \) depends on the censoring
distribution and is typically not of direct interest in
itself. In the following, we describe how an estimate of
\( F \), as provided by the state learner, can be used to
estimate more relevant parameters of interest, such as, e.g.,
survival probabilities.

@@

Event-free survival probabilities and risk predictions can be
obtained from the state-occupation function \( F \) under the
assumption of conditional independent censoring and
positivity, as introduced in Section~\ref{sec:framework}. By
equations~(\ref{eq:lambdaj}) and~(\ref{eq:gamma}) and the
definition of \( F \), we have
\begin{equation}
  \label{eq:7}
  % \Gamma(t , x) = \int_0^t \frac{F(\diff s, -1, x )}{F(s-,
  % 0, x )}, \quad \text{and} \quad
  \Lambda_j(t , x) 
  = \int_0^t  \frac{F(\diff s, j, x )}{F(s-, 0, x )},
  \quad j \in \{1,2\}.
\end{equation}
Equation~\eqref{eq:surv-def} provides a formula for obtaining
event-free survival probabilities from the cause-specific
cumulative hazard functions. Similarly, cause-specific risk
predictions can be obtained from \( \Lambda_1 \) and
$\Lambda_2$ using the formula
\citep[e.g.,][]{benichou1990estimates,
  ozenne2017riskregression},
\begin{equation}
  \label{eq:cs-risk-def}
  Q(T \leq t, D = j \mid X=x) =
  \int_0^t \exp\left\{-\Lambda_{1}(u \mid x)-\Lambda_{2}(u
    \mid x)\right\}  \Lambda_j(\diff u \mid x),
  \quad j \in \{1,2\}.
\end{equation}
Hence, given the state learner's estimate of \( F \) we can
use equation~\eqref{eq:7} to obtain estimates of the
cause-specific cumulative hazard functions $\Lambda_j$, which
can in turn be used to obtain estimates of the event-free
survival probabilites and cause-specific risks through
equations~\eqref{eq:surv-def} and~\eqref{eq:cs-risk-def}.

@@

The framework of \textit{targeted} or \textit{debiased}
machine learning is a general methodology for combining
data-adaptive modeling, such as machine and super learning,
with asymptotically valid statistical inference for
interpretable target parameters \citep{van2006targeted,
  chernozhukov2018double}. \fxnote*{add more references?}{This
  framework has been applied extensively for causal inference
  based on observational data
  \citep[e.g.,][]{van2011targeted,kennedy2016semiparametric,rytgaard2022targeted}.}
A targeted or debiased estimator is constructed by first
estimation high-dimensional nuisance parameter using machine
learning, and then employing a debiasing or targeting step
using tools from semi-parametric efficiency theory
\citep{pfanzagl1985contributions,bickel1993efficient,van2003unified,tsiatis2007semiparametric,kennedy2022semiparametric}.
In the context of right-censored competing risk data, the
high-dimensional nuisance parameters needed to be estimated
will typically include the cause-specific cumulative hazard
functions and the cumulative hazard function for censoring
\citep[e.g.,][]{van2003unified,rytgaard2021estimation,rytgaard2022targeted}.
The state learner is hence particularly well suited for
targeted and debiased machine learning as it immediately
provides estimates of these nuisance parameters. While it was
our original motivation for developing the state learning, we
leave the study of the state learner in the context of
targeted and debiased machine learning for a future paper.

@@

For all super learners we use a library consisting of three
learners: The Kaplan-Meier estimator
\citep{kaplan1958nonparametric,Gerds_2019prodlim}, a Cox model
with main effects \citep{cox1972regression, survival-package},
and a random survival forest
\citep{ishwaran2008random,rfsrc-paclage}. We use the same
library to learn the outcome distribution and the censoring
distribution. Note that the three learners in our library of
learners can be used to learn the cumulative hazard functions
of the outcome and the censoring distribution. The latter
works by training the learner on the data set \( \data_n^c \),
where \( \data_n^c = \{O_i^c\}_{i=1}^n \) with
\( O_i^c = (\tilde{T}_i, 1-\Delta_i, X_i) \). When we say that
we use a learner for the cumulative hazard function of the
outcome to learn the cumulative hazard function of the
censoring time, we mean that the learner is trained on
\( \data_n^c \).
